Unity in diversity: Children from 17 diverse communities process gaze in similar ways

### 2024-03-19 paper/paper.Rmd
clean repo

    affiliation   : "2,10"

@@

  - name          : "Daniel B. M. Haun"

@@

affiliation:
  - id            : "1"
    institution   : "Institute of Psychology in Education, Leuphana University Lüneburg"
  - id            : "2"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology"
  - id            : "3"
    institution   : "Budongo Conservation Field Station"
  - id            : "4"
    institution   : "School of Psychology, University of Auckland"
  - id            : "5"
    institution   : "Université Marien Ngouabi"
  - id            : "6"
    institution   : "Department of Psychology, Stanford University"
  - id            : "7"
    institution   : "School of Psychology, University of Plymouth"
  - id            : "8"
    institution   : "Department of Psychology, Koç University"
  - id            : "9"
    institution   : "Division of Psychology, Communication, and Human Neuroscience, University of Manchester"   
  - id            : "10"
    institution   : "Department of Psychology and Social Work, University of Namibia"
  - id            : "11"
    institution   : "Department of Psychology, University of York"
  - id            : "12"
    institution   : "CAS Key Laboratory of Behavioral Science, Institute of Psychology, Chinese Academy of Sciences"
  - id            : "*"
    institution   : "joint first author"

authornote: |
   The authors would like to thank Luke Maurits for statistical advice. Manuel Bohn was supported by a Jacobs Foundation Research Fellowship (2022-1484-00). We are gratful to thank all children and caregivers for participating in the study.

abstract: |
  Theoretical accounts assume that key features of human social cognition are universal. Here we focus on gaze following, the bedrock of social interactions and coordinated activities, to test this claim. In this comprehensive cross-cultural study spanning five continents and 17 distinct cultural communities, we examined the development of gaze following in early childhood. We identified key processing signatures through a computational model that assumes that participants follow an individual’s gaze by estimating a vector emanating from the eye-center through the pupil. Using a single reliable touchscreen-based task, we found these signatures in all communities, suggesting that children worldwide processed gaze in highly similar ways. Additionally, we found a cross-culturally consistent relationship between children’s exposure to touchscreens and their performance in the task, which also explained absolute differences between communities. These results provide strong evidence for a universal process underlying a foundational socio-cognitive ability in humans that can be reliably inferred even in the presence of cultural variation in overt behavior.

@@

Human socio-cognitive skills enable unique forms of communication and cooperation that provide a bedrock for cumulative culture and the formation of complex societies  [@tomasello2020adaptive; @laland2021understanding; @wellman2014making; @henrich2016secret; @tomasello2003makes; @legare2019development; @heyes2018cognitive]. The eyes are the proverbial "window to the mind" and eye gaze is essential for many social reasoning processes  [@shepherd2010following; @doherty2006development]. Others’ eye gaze is used to infer their focus of visual attention, which is a critical aspect of coordinated activities, including communication and cooperation [@tomasello2007reliance; @scaife1975capacity]. The ability to follow gaze emerges early in childhood [@tang2023slow; @gredeback2010development; @byers2021development; @del2019developmental] and individual differences in children’s gaze-following ability predict later life outcomes, most notably, communicative abilities [@brooks2005development; @carpenter1998social]. Difficulties with gaze following have been linked to developmental disorders, including Autism [@itier2009neural].

There is a widely-held assumption, that, despite substantial variation in developmental contexts, gaze-following works and develops in the same way across human societies [@tomasello2019becoming]. However, this view is seemingly at odds with the substantial diversity in socio-cognitive development that cross-cultural studies have revealed [@miller2018contributions; @wellman2014making; @mayer2013synchrony; @dixson2018scaling; @taumoepeau2019cross]. One potential source for this paradox lies in the reliance on aggregated measures in cross-cultural studies. Absolute differences in mean performance across communities are interpreted as a signal of different underlying cognitive processes. In this study, we resolve this paradox by instead focusing on processing signatures that can be investigated independently of absolute community-level differences. This allows us to directly evaluate the empirical foundation of claims about universal features of human social cognition. To this end, we conducted a pre-registered, large-scale, cross-cultural study on the development of gaze-following abilities to study potentially universal processing signatures. These signatures were derived from a computational model that assumes that participants follow gaze by estimating a vector emanating from the eye center through the pupil [@prein2023variation].

The `r n_distinct(data$subjid)` participants who took part in the study lived in 17 different communities across 14 countries and five continents (Fig. \@ref(fig:fig1)A, Tab. \@ref(tab:tab1)). These countries represent ~46% of the world’s population. Communities covered a broad spectrum of geographical locations, social and political systems, languages, and subsistence styles (see Supplemental Materials). This diversity allowed us to overcome the common pitfall of cross-cultural studies that largely compared urban communities from the global north to rural communities from the global south [@barrett2020towards].

```{r fig1, include = T, fig.align = "center", fig.cap = "(A) Data collection sites. Points show the approximate geographical location of the data collection sites, coloring shows the sample sizes. (B) Screenshot from the task. The scene depicts the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and click on the location on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adapted to each cultural setting. (C) Drawings used as agents across cultural settings. (D) Graphical illustration of the cognitive model. Individuals infer the target of an agent’s attention by estimating a vector based on the position of the pupils within the eyes. This process is noisy, illustrated by the different vectors (transparent lines). Individuals differ in their level of precision (indicated by sigma). For a given level of precision, the closer the  target lands to the left or the right, the less precise the model predicts individuals to be. Solid and transparent dots show simulated means and individual data points to illustrate the predicted effect of target position.", out.width="75%"}
knitr::include_graphics("../visuals/fig1.png")
```

```{r tab1, include = T}
tba1 <- data %>%
  separate(community, into = c("community", "country"), sep = " \\(")%>%
  mutate(country = str_remove(country, "\\("),
         country = str_remove(country, "\\)"))%>%
  distinct(subjid, .keep_all = T)%>%
  group_by(community, country, continent)%>%
  summarise(N = n_distinct(subjid), 
            m = sum(sex == "m", na.rm = T), 
            mean_age = mean(ageinyears),
            min_age = min(ageinyears), 
            max_age = max(ageinyears), 
            touchscreen = mean(touchscreen, na.rm = T))%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  mutate(age = paste0(mean_age, " (", min_age, " - ", max_age, ")"))%>%
  mutate(N = paste0(N, " (", m,")"))%>%
  select(-c(mean_age, min_age, max_age))%>%
  arrange(continent, country)%>%
  group_by(continent) %>%
  mutate(continent = as.character(continent),
         continent = replace(continent, duplicated(continent), " ")) %>%
  ungroup() %>%
  mutate(country = replace(country, duplicated(country), " "))%>%
  mutate(language = c("Spanish (Rioplatense)", "Spanish (Mexican)", "English (American)", "Hai||om", "Khwedam", "English (Nigerian)*", "English (Nigerian)", "BaYaka", "Lingala", "Swahili", "Bemba", "German", "English (British)", "Mandarin", "English (Indian) / Marathi", "Turkish", "English (New Zealand)"),
           market_integration = c("high", "medium", "high", "low", "low", "high", "high", "low", "low", "medium", "medium", "high", "high", "high", "high", "high", "high"))%>%
  select(continent, country, community, N, age, language, market_integration, touchscreen)

apa_table(tba1,
          col.names = linebreak(c("Continent", "Country", "Community", "N(male)", "Age (range)", "Language", "Market integration","Touchscreen")),
          caption = "Participant demographics.",
          #align = paste0("m{", 1/(ncol(tba1) + 1), "\\linewidth}"),
          longtable = TRUE,
          note = "*Local collaborators and piloting suggested that Nigerian English is suitable for Windhoek as well.",
          escape=TRUE)
```

We used an animated picture book tablet task in which participants had to locate a hidden object based on observing an agent’s gaze. Children watched a balloon disappear behind a hedge. An agent followed the trajectory of the balloon with their eyes (Fig. \@ref(fig:fig1)B). The key dependent variable was the (im)precision with which children located the agent’s focus of attention, that is, the deviation between where the agent looked (where the object was) and the child’s response. We adapted visuals and audio instructions specifically to each of the 17 communities. Previous work demonstrated excellent individual-level measurement properties for this task in a German sample [@prein2023tango].

@@

Nevertheless, there were also marked differences between communities (see Fig. \@ref(fig:fig2)A). In a six-fold cross-validation procedure, we trained a regression model on a subset of the data (training data) to later predict the held-out data (testing data) [@stengelin2023children]. We found that a model assuming cross-cultural variation in average performance as well as cross-cultural variation in developmental trajectories outperformed simpler models – assuming no variation in the shape of developmental trajectories or nor variation between settings at all – in `r round(mean(cv_cult$m3_beats_m2),2)*100`% of cases (see Supplemental Material). At first glance, it seems that highly market-integrated communities around the globe showed higher levels of precision compared to less market-integrated communities (compare Tab. \@ref(tab:tab1) and Fig. \@ref(fig:fig2)A). However, numerous alternative groupings (e.g., average levels of education or average household size) are possible and, instead, these results could be best understood in terms of exposure to touch-screen devices; a finding we discuss in more detail below. Importantly, average differences in precision between communities were small compared to differences between individuals: communities did not form homogeneous clusters but largely overlapping distributions in that some individuals from communities with a lower average level of precision performed better compared to some individuals from a setting with a very high average level of precision. Similarly, in all communities, some 4-year-olds outperformed children two years older than them (see Fig. \@ref(fig:fig2)A). The lack of adequate individual-level measurement instruments  in previous large-scale developmental cross-cultural studies made it impossible to contrast these perspectives. The substantial overlap between communities found here speaks against categorical differences in gaze following and is suggestive of a similar underlying process. However, consistent developmental improvements alone cannot inform us about the cognitive processes children use when locating the agent’s focus of attention.

```{r fig2, include = T, fig.align = "center", fig.cap = "A) Developmental trajectory across and B) by community. The developmental trajectories are predicted based on a model of the data aggregated for each participant. C) Performance by screen section across and D) by community. Each bin covers 1/10th of the screen. Points show means, and error bars 95% confidence intervals for the data within that bin aggregated across participants. Transparent dots in A) and C) show aggregated data for each individual.", out.width="100%"}
knitr::include_graphics("../visuals/pvis_pred.png")
```

Recent computational work modeled gaze following as social vector estimation [@prein2023variation]. When observing the eyes, onlookers estimate a vector emanating from the center of the eye that runs through the pupil. The focus of attention is the location where the estimated vectors from both eyes hit a surface (Fig. \@ref(fig:fig1)D). It is assumed that this estimation process has some uncertainty and that individuals vary in their level of uncertainty. As a consequence, even though individuals use the same general process, they might differ in their absolute levels of precision. Crucially, this process model predicts a clear performance signature in our gaze following task: Trials in which the agent looks further away from the center (i.e. to the left or right side of the screen) should result in lower levels of precision compared to trials in which the agent looks closer to the middle. This prediction is best understood by considering a similar phenomenon: pointing a torch light to a flat surface. The width of the light beam represents each individual’s level of uncertainty in vector estimation. When the torch is directed straight down, the light beam is concentrated in a relatively small area. When the torch is rotated to the side, the light from one half of the cone must travel further than the light from the other half to reach the surface. As a consequence, the light is spread over a wider area (see Fig. \@ref(fig:fig1)D).

This processing signature was clearly visible across all 17 communities. Precision decreased when the agent looked at locations further away from the center (fixed effect: $\beta$ = `r post_fix%>%filter(term == "targetcentralityx")%>%pull(Estimate)`, 95% HDI (`r post_fix%>%filter(term == "targetcentralityx")%>%pull(Q2.5)` - `r post_fix%>%filter(term == "targetcentralityx")%>%pull(Q97.5)`); range of community-level effects: $\beta_{min}$ = `r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%tail(1)%>%pull(mean)`, 95% HDI (`r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%tail(1)%>%pull(lci)` - `r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%tail(1)%>%pull(uci)`) to $\beta_{max}$ = `r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%head(1)%>%pull(mean)`, 95% HDI (`r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%head(1)%>%pull(lci)` - `r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%head(1)%>%pull(uci)`). Visualization of the data showed the predicted u-shaped pattern in all communities (see Fig. \@ref(fig:fig2)B). These results indicate a universal cognitive process used by children in all communities. There are, however, alternative ways in which the u-shaped pattern might arise: if participants ignored the agent’s gaze and instead always selected the middle of the screen (center bias) or randomly select locations (random guessing), precision would also decrease when the balloon lands further away from the center. To rule out these alternative explanations, we compared three cognitive models that made different assumptions about how participants’ responses were generated: the focal vector-based gaze estimation model described above, a center-bias model where participants always select the center, and a random guessing model where participants select random locations. For every community, we found overwhelming support for the gaze estimation model (min $BF_{10}$ > 100 000 for comparisons with both alternative models). Taken together, children from all 17 communities processed gaze in similar ways.

@@

However, there was substantial variation between individuals that could not be explained by differential exposures to touch screens alone. For example, in Malatya (Türkyie) where 100% of children had access to touch screens there was still substantial variation between individuals (see Fig. 1B).  This strongly indicates that other factors likely contributed to individual differences. Social interaction has been highlighted as an important driver of social-cognitive development [e.g., @tomasello2019becoming; @carpendale2020makes; @barresi1996intentional; @perner1994theory; @rakoczy2022foundations] and thus we hypothesized (and pre-registered) that more opportunities for social interaction -- approximated by living in larger households with more children -- would be associated with higher levels of precision. At first glance, the opposite seemed to be the case: there was a substantial zero-order correlation between the number of children living in the household and average imprecision (`r round(ctest$estimate,2)`, 95% CI = `r round(ctest$conf.int[1],2)` - `r round(ctest$conf.int[2],2)`) suggesting that having more opportunities for social interactions was related to poorer performance. However, this correlation is spurious and reflects a confounding of household composition with exposure to technology on a community level. When predicting performance by relative opportunities for social interactions within a community -- while accounting for absolute differences and the prevalence of touchscreens -- we found no strong associations between any of the demographic indicators and performance (see Supplemental Material). The reason for an absence of such an association may be a lack of resolution: household composition is very far removed from the factors that previous work has suggested to be related to the development of gaze following in younger children, such as attachment quality or the use of gaze in early communicative interactions [@astor2020social; @movellan2002development; @senju2015early]. Future work could increase the resolution with which everyday experiences in children from diverse communities are recorded to compare the drivers behind development as we observe it. Recent work in the field of language acquisition has shown how technological innovations allowed for direct recording of social interactions across communities which can be used to close this explanatory gap [@bergelson2023everyday; @donnelly2021longitudinal].

@@

The cognitive processes underlying gaze following might be rooted in humans’ evolved cognitive architecture, which is -- presumably -- later refined during social (i.e., cultural) interaction [@astor2020social; @movellan2002development; @senju2015early]. The phylogenetic roots of these processes might possibly lie much deeper as primates from a wide range of species follow gaze [@rosati2009looking; @tomasello1998five; @itakura2004gaze; @kano2014cross]. Yet, similarities in overt behavior do not imply the same underlying cognitive processes. The present study defines clear performance signatures that can be explored in other species to test such evolutionary hypotheses.

>>>>

    affiliation   : "2,10,+"

@@

  - name          : "Daniel B. M. Haun (ORCID: 0000-0002-3262-645X)"
    affiliation   : "2,+"

@@

affiliation:
  - id            : "1"
    institution   : "Institute of Psychology in Education, Leuphana University Lüneburg"
  - id            : "2"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology"
  - id            : "3"
    institution   : "Budongo Conservation Field Station"
  - id            : "4"
    institution   : "School of Psychology, University of Auckland"
  - id            : "5"
    institution   : "Université Marien Ngouabi"
  - id            : "6"
    institution   : "Department of Psychology, Stanford University"
  - id            : "7"
    institution   : "School of Psychology, University of Plymouth"
  - id            : "8"
    institution   : "Department of Psychology, Koç University"
  - id            : "9"
    institution   : "Division of Psychology, Communication, and Human Neuroscience, University of Manchester"   
  - id            : "10"
    institution   : "Department of Psychology and Social Work, University of Namibia"
  - id            : "11"
    institution   : "Department of Psychology, University of York"
  - id            : "12"
    institution   : "CAS Key Laboratory of Behavioral Science, Institute of Psychology, Chinese Academy of Sciences"
  - id            : "*"
    institution   : "joint first author"
  - id            : "+"
    institution   : "joint last author"

authornote: |
   The authors would like to thank Luke Maurits for statistical advice. Manuel Bohn was supported by a Jacobs Foundation Research Fellowship (2022-1484-00). We are gratful to thank all children and caregivers for participating in the study. We thank the Max Planck Society for the Advancement fo Science.

abstract: |
  Theoretical accounts assume that key features of human social cognition are universal. Here we focus on gaze following, the bedrock of social interactions and coordinated activities, to test this claim. In this comprehensive cross-cultural study spanning five continents and 17 distinct cultural communities, we examined the development of gaze following in early childhood. We identified key processing signatures through a computational model that assumes that participants follow an individual’s gaze by estimating a vector emanating from the eye-center through the pupil. Using a single reliable touchscreen-based task, we found these signatures in all communities, suggesting that children worldwide processed gaze in highly similar ways. Absolute differences in performance between groups are accounted for by a cross-culturally consistent relationship between children’s exposure to touchscreens and their performance in the task. These results provide strong evidence for a universal process underlying a foundational socio-cognitive ability in humans that can be reliably inferred even in the presence of cultural variation in overt behavior.

@@

Human socio-cognitive skills enable unique forms of communication and cooperation that provide a bedrock for cumulative culture and the formation of complex societies  [@tomasello2020adaptive; @laland2021understanding; @wellman2014making; @henrich2016secret; @tomasello2003makes; @legare2019development; @heyes2018cognitive]. The eyes are the proverbial "window to the mind" and eye gaze is essential for many social reasoning processes  [@shepherd2010following; @doherty2006development]. Others’ eye gaze is used to infer their focus of visual attention, which is a critical aspect of coordinated activities, including communication and cooperation [@tomasello2007reliance; @scaife1975capacity]. The ability to follow gaze emerges early in childhood [@tang2023slow; @gredeback2010development; @byers2021development; @del2019developmental], and individual differences in children’s gaze-following ability predict later life outcomes, most notably communicative abilities [@brooks2005development; @carpenter1998social]. Difficulties with gaze following have been linked to developmental disorders, including Autism [@itier2009neural].

There is a widely held assumption that, despite substantial variation in developmental contexts, gaze-following works and develops in the same way across human societies [@tomasello2019becoming]. However, this view is seemingly at odds with the substantial diversity in socio-cognitive development that cross-cultural studies have revealed [@miller2018contributions; @wellman2014making; @mayer2013synchrony; @dixson2018scaling; @taumoepeau2019cross]. One potential source for this paradox lies in the reliance on aggregated measures in cross-cultural studies. Absolute differences in mean performance across communities are interpreted as a signal of different underlying cognitive processes. In this study, we resolve this paradox by instead focusing on processing signatures that can be investigated independently of absolute community-level differences. This allows us to directly evaluate the empirical foundation of claims about universal features of human social cognition. To this end, we conducted a pre-registered, large-scale, cross-cultural study on the development of gaze-following abilities to study potentially universal processing signatures. These signatures were derived from a computational model that assumes that participants follow gaze by estimating a vector emanating from the eye center through the pupil [@prein2023variation].

The `r n_distinct(data$subjid)` participants who took part in the study lived in 17 different communities across 14 countries and five continents (Fig. \@ref(fig:fig1)A, Tab. \@ref(tab:tab1)). These countries represent ~46% of the world’s population. Communities covered a broad spectrum of geographical locations, social and political systems, languages, and subsistence styles (see Supplemental Materials). This diversity allowed us to overcome the common pitfall of cross-cultural studies that compare urban communities from the global north to rural communities from the global south [@barrett2020towards].

```{r fig1, include = T, fig.align = "center", fig.cap = "(A) Data collection sites. Points show the approximate geographical location of the data collection sites, coloring shows the sample sizes. (B) Screenshots from the task. Screenshots from the task. The upper scene depicts the start and the lower the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and touch the location on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adapted to each cultural setting. (C) Drawings used as agents across cultural settings. (D) Graphical illustration of the cognitive model. Individuals infer the target of an agent’s attention by estimating a vector based on the position of the pupils within the eyes. This process is noisy, illustrated by the different vectors (transparent lines). Individuals differ in their level of precision (indicated by sigma). For a given level of precision, the further the target lands from the centre of the screen, the less precise the model predicts individuals to be. Solid and transparent dots show simulated means and individual data points to illustrate the predicted effect of target position.", out.width="75%"}
knitr::include_graphics("../figures/fig1.png")
```

```{r tab1, include = T}
tba1 <- data %>%
  separate(community, into = c("community", "country"), sep = " \\(")%>%
  mutate(country = str_remove(country, "\\("),
         country = str_remove(country, "\\)"))%>%
  distinct(subjid, .keep_all = T)%>%
  group_by(community, country, continent)%>%
  summarise(N = n_distinct(subjid), 
            m = sum(sex == "m", na.rm = T), 
            mean_age = mean(ageinyears),
            min_age = min(ageinyears), 
            max_age = max(ageinyears), 
            touchscreen = mean(touchscreen, na.rm = T))%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  mutate(age = paste0(mean_age, " (", min_age, " - ", max_age, ")"))%>%
  mutate(N = paste0(N, " (", m,")"))%>%
  select(-c(mean_age, min_age, max_age))%>%
  arrange(continent, country)%>%
  group_by(continent) %>%
  mutate(continent = as.character(continent),
         continent = replace(continent, duplicated(continent), " ")) %>%
  ungroup() %>%
  mutate(country = replace(country, duplicated(country), " "))%>%
  mutate(language = c("Spanish (Rioplatense)", "Spanish (Mexican)", "English (American)", "Hai||om", "Khwedam", "English (Nigerian)2", "English (Nigerian)", "BaYaka", "Lingala", "Kiswahili", "Bemba", "German", "English (British)", "Mandarin", "English (Indian) / Marathi", "Turkish", "English (New Zealand)"))%>%
  select(continent, country, community, N, age, language, touchscreen)

apa_table(tba1,
          col.names = linebreak(c("Continent", "Country", "Community", "N(male)", "Age (range)", "Language", "Touchscreen exposure1")),
          caption = "Participant demographics.",
          #align = paste0("m{", 1/(ncol(tba1) + 1), "\\linewidth}"),
          longtable = TRUE,
          note = "1 Proportion of participants who have access to touchscreens according to parental questionnaire. 2 Local collaborators and piloting suggested that Nigerian English is suitable for Windhoek as well.",
          escape=TRUE)
```

We used an animated picture book tablet task in which participants had to locate a hidden object based on observing an agent’s gaze. Children watched a balloon disappear behind a hedge. An agent followed the trajectory of the balloon with their eyes (Fig. \@ref(fig:fig1)B). The key dependent variable was the (im)precision with which children located the agent’s focus of attention, that is, the deviation between where the agent looked (where the object was) and the child’s response. We adapted visuals and audio instructions specifically for each of the 17 communities. Previous work demonstrated excellent individual-level measurement properties for this task in a German sample [@prein2023tango].

@@

Nevertheless, there were also marked differences between communities (see Fig. \@ref(fig:fig2)A). In a six-fold cross-validation procedure, we trained a regression model on a subset of the data (training data) to later predict the held-out data (testing data) [@stengelin2023children]. This procedure was repeated 100 times. We found that a model assuming cross-cultural variation in average performance as well as cross-cultural variation in developmental trajectories outperformed simpler models – assuming no variation in the shape of developmental trajectories or nor variation between settings at all – in `r round(mean(cv_cult$m3_beats_m2),2)*100`% of cases (see Supplemental Material). There are numerous ways in which communities could be grouped that would fall in line with these absolute differences (e.g., market integration, average levels of education, or average household size). However, we deliberately want to avoid any such overly simplistic explanation. Instead, we think these results can be best understood in terms of exposure to touch-screen devices;; a finding we discuss in more detail below. Importantly, average differences in precision between communities were small compared to differences between individuals: communities did not form homogeneous clusters but largely overlapping distributions in that some individuals from communities with a lower average level of precision performed better compared to some individuals from a setting with a very high average level of precision. Similarly, in all communities, some 4-year-olds outperformed children two years older than them (see Fig. \@ref(fig:fig2)A). The lack of adequate individual-level measurement instruments  in previous large-scale developmental cross-cultural studies made it impossible to contrast these perspectives. The substantial overlap between communities found here speaks against categorical differences in gaze following and is suggestive of a universal underlying process. However, consistent developmental improvements alone cannot inform us about the cognitive processes children use when locating the agent’s focus of attention.

```{r fig2, include = T, fig.align = "center", fig.cap = "A) Developmental trajectory across and B) by community. The developmental trajectories are predicted based on a model of the data aggregated for each participant. C) Performance by target location on the screen across, and D) by community. Each bin covers 1/10th of the screen. Points show means, and error bars 95% confidence intervals for the data within that bin aggregated across participants. Transparent dots in A) and C) show aggregated data for each individual.", out.width="100%"}
knitr::include_graphics("../figures/pvis_pred.png")
```

Recent computational work modeled gaze following as social vector estimation [@prein2023variation]. When observing the eyes, onlookers estimate a vector emanating from the center of the eye that runs through the pupil. The focus of attention is the location where the estimated vectors from both eyes hit a surface (Fig. \@ref(fig:fig1)D). It is assumed that this estimation process has some uncertainty and that individuals vary in their level of uncertainty. As a consequence, even though individuals use the same general process, they might differ in their absolute levels of precision. Crucially, this process model predicts a clear performance signature in our gaze following task: Trials in which the agent looks further away from the center should result in lower levels of precision compared to trials in which the agent looks closer to the center. This prediction is best understood by considering a similar phenomenon: pointing a torch light to a flat surface. The width of the light beam represents each individual’s level of uncertainty in vector estimation. When the torch is directed straight down, the light beam is concentrated in a relatively small area. When the torch is rotated to the side, the light from one half of the cone must travel further than the light from the other half to reach the surface. As a consequence, the light is spread over a wider area (see Fig. \@ref(fig:fig1)D).

This processing signature was clearly visible across all 17 communities. Precision decreased when the agent looked at locations further away from the center (fixed effect: $\beta$ = `r post_fix%>%filter(term == "targetcentralityx")%>%pull(Estimate)`, 95% HDI (`r post_fix%>%filter(term == "targetcentralityx")%>%pull(Q2.5)` - `r post_fix%>%filter(term == "targetcentralityx")%>%pull(Q97.5)`); range of community-level effects: $\beta_{min}$ = `r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%tail(1)%>%pull(mean)`, 95% HDI (`r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%tail(1)%>%pull(lci)` - `r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%tail(1)%>%pull(uci)`) to $\beta_{max}$ = `r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%head(1)%>%pull(mean)`, 95% HDI (`r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%head(1)%>%pull(lci)` - `r post_agg%>%filter(type == "targetcentrality")%>%arrange(mean)%>%head(1)%>%pull(uci)`). Visualization of the data showed the predicted u-shaped pattern in all communities (see Fig. \@ref(fig:fig2)B). These results indicate a universal cognitive process used by children in all communities. There are, however, alternative ways in which the u-shaped pattern might arise: if participants ignored the agent’s gaze and instead always selected the middle of the screen (center bias) or randomly selected locations (random guessing), precision would also decrease when the balloon lands further away from the center. To rule out these alternative explanations, we compared three cognitive models that made different assumptions about how participants’ responses were generated: the focal vector-based gaze estimation model described above, a center-bias model where participants always select the center, and a random guessing model where participants select random locations. For every community, we found overwhelming support for the gaze estimation model (min $BF_{10}$ > 100 000 for comparisons with both alternative models). Taken together, children from all 17 communities processed gaze in similar ways.

@@

However, there was substantial variation between individuals that could not be explained by differential exposures to touch screens alone. For example, in Malatya (Türkyie) where 100% of children had access to touchscreens there was still substantial variation between individuals (see Fig. 1B).  This strongly indicates that other factors likely contributed to individual differences. Social interaction has been highlighted as an important driver of social-cognitive development [e.g., @tomasello2019becoming; @carpendale2020makes; @barresi1996intentional; @perner1994theory; @rakoczy2022foundations] and thus we hypothesized (and pre-registered) that more opportunities for social interaction -- approximated by living in larger households with more children -- would be associated with higher levels of precision. When predicting performance by relative opportunities for social interactions within a community -- while accounting for absolute differences and the prevalence of touchscreens -- we found no strong associations between any of the demographic indicators and performance (see Supplemental Material). Whilst household size was a useful proxy for regular social interaction opportunities, the measure does not directly measure the factors that previous work has suggested to be related to the development of gaze following in younger children, such as attachment quality or the use of gaze in early communicative interactions [@astor2020social; @movellan2002development; @senju2015early]. Future work could increase the resolution with which everyday experiences in children from diverse communities are recorded to compare the drivers behind social-cognitive development as we observe it. Recent work in the field of language acquisition has shown how technological innovations allowed for direct recording of social interactions across communities which can be used to close this explanatory gap [@bergelson2023everyday; @donnelly2021longitudinal].

@@

The cognitive processes underlying gaze following might be rooted in humans’ evolved cognitive architecture, which is -- presumably -- later refined during social interaction [@astor2020social; @movellan2002development; @senju2015early]. The phylogenetic roots of these processes might possibly lie much deeper as primates from a wide range of species follow gaze [@rosati2009looking; @tomasello1998five; @itakura2004gaze; @kano2014cross]. Yet, similarities in overt behavior do not imply the same underlying cognitive processes. The present study defines clear performance signatures that can be explored in other species to test such evolutionary hypotheses.
